Secure Multi-Party Computation with SCALE-MAMBA

### 2019-04-21 paper.tex
Sun Morn

\section{Motivation and Introduction}\label{sec:intromoti}
In today's world, and even more in the future, there exist a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all of it. Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive or it has a high business value.\\
A basic example is an anonymous voting. In most elections there is a third party that has to aggregate and count the votes. For instance, in Germany's elections[TODO: LINK] this is done by specific people counting a small amount of votes and then forwarding the information to a central institution. The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time. So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of voters.\\
Another use case is the exchange of medical data. Medical data of patients is highly confidential and should never be published. But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it would be helpful to compare the symptoms to the diagnoses of other doctors.\\
Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
The basic problem is the evaluation of a common function, to which multiple parties have different inputs, without revealing these inputs. Although there are a lot of theoretical solutions for this problem, real-world implementations are still not performing on the desired level. So in the last years the research shifted more into the area of how to improve the performance of existing solutions while still guaranteeing a desired level of security.\\
This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven. This system is one of the forerunners in SMPC and combines a lot of research by different teams to provide a platform with highly performant SMPC. Nevertheless, this paper will at first give a short introduction into the history of SMPC and will try to explain its basic technologies to the reader. It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially taking a look on which different requirements are fulfilled by each. Ultimately, it will try to give an outlook what might be coming in the next years.\\

\section{Cryptographic Basics}\label{sec:basics}

Before delving deeper into SMPC, this section will shortly recap on cryptography basics that might be unknown to the reader. These basics are not solely used by SMPC, but nevertheless, play a huge role in a lot of protocols.

\subsection{Message Authentification Codes}
Message Authentification Codes (MACs) are used to guarantee the integrity and authenticity of a message. Basically, a MAC is a key that is used to generate a tag for a message. This tag is used to verify the identity of the sender and that the content of the message was not changed.\\
For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
\begin{itemize}
\item The MAC algorithm \(\text{MAC}(K,m)\rightarrow t\) takes a key \(K\) and a message \(m\) and outputs a tag \(t\)\\
\item The verficiation algorithm \(\text{VER}(K,m,t)\rightarrow 0/1\) takes a key \(K\), a message \(m\) and a tag \(t\) and outputs accept(\(1\)) or reject(\(0\))\\
\end{itemize}
The correctness can be easily seen: \(\forall K: \text{VER}(K,m,\text{MAC}(K,m))=1\)\\

\subsection{Homomorphic Encryption}
Encryption as a basic concept is simple. Similiar to MACs, a tupel (E,D) of algorithms is needed.\\
\begin{itemize}
\item The enncryption algorithm \(\text{E}(K,m)\rightarrow c\) takes a key \(K\) and a message \(m\) and outputs a cyphertext \(t\) that is the encryption of the message \(m\)\\
\item The decryption algorithm \(\text{D}(K,c)\rightarrow m\) takes a key \(K\) and a cyphertext \(m\) and outputs the decrypted message \(m\), assuming c was not modified\\
\end{itemize}
Alternatively, in a public key setting, the key for encryption is public and there is a different key for decryption that is privat.\\
The problem is, that modifying the cypertext in any way makes the decryption useless. In some cases it is desirable do do computations on cypertexts, because then it is possible to make calculations without revealing the unencrypted data. That is why Homomorphic Encryption was developed. Let's assume that we have a ring \((R,+,*)\) as our message space. Fully Homomorphic Encryption (FHE) is preserving the ring structure. Therefore, it is possible to do additions and multiplications of cypertexts and when decrypting the result it is the same value as when doing the same additions and multiplications on the real values.\\
Althought FHE has amazing properties for a lot of usecases, in praxis it is way too slow. Because of all the overhead generated by the encryption scheme to guarantee the preservation of the ring structure, the current algorithms are not feasible for most applications.

\subsection{Somewhat Homomorphic Encryption}
That is where Somewhat Homomorphic Encryption (SHE) becomes relevant. SHE only supports one operation, and in most cases that is addition. This takes away a lot of possibilities but, therefore, the performance gain is immense. So immense, that SHE is used a lot in practice, for example in SCALE-MAMBA.\\
The SHE scheme used in SCALE-MAMBA, is based on the so called Brakerski-Gentry-Vaikuntanathan cryptosystem(BGV).\\

@@

\subsection{Beginnings}
As we saw in the introduction, there are a lot current and much more upcoming use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret. The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yao's millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their respective wealth). \\
These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\

\subsection{Problem statement}
Formally, the SMPC problem can be stated as follows:
\begin{itemize}
\item Given a finite field \( \mathcal{F}_p=\{0,...,p-1\}\) with a prime p
\item Given \(n\) parties \(P_1,...,P_n\) participating in the computation
\item Given \(m\) secret inputs \(x_1,...,x_m\) with each of them hold by one of the parties
\item Given a function \(f(x_1,...,x_m)\)
\item Goal: Evaluation of the function \(f(x_1,...,x_m)\) while guaranteeing secrecy and correctness
\end{itemize}
\textbf{Secrecy:} If party \(P_i\) holds secret \(x_j\), then at the end of the evaluation no information regarding \(x_j\) is revealed to the other parties.\\
\textbf{Correctness:} At the end of the computation, the result is correct.\\\\
In a lot of cases each party has exactly one corresponding secret input, but this is not necessarily required. There can also be parties without input and parties with multiple inputs.\\
Now we will take a short look on two of the first protocols, which both are the basis for a whole family of SMPC schemes.

\subsection{Yao's Garbled Circuit}
There are typically two families, into which most of the SMPC protocols can be categorized. One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}. The basic idea is that, in a two party setting, one party, Alice, garbles the Boolean circuit that has to be evaluated and sends it to the second party, Bob. Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an encrypted way. At the end they exchange the results and can compute the result this way.

\subsection{Shamirs Secret Sharing Scheme}
This paper will not further discuss the family of garbled circuit protocols, but rather the second family which are protocols based on so called Linear Secret Sharing. 
These type of protocols were introduced by Adi Shamir at a similar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
The basic idea of this approach is that multiple parties split their secrets into so called shares and then distribute them to the other parties. In Shamir's scheme this is done by generating random polynomials with a degree depending on the security level and then calculating the shares as points in this polynomial. Computations are executed on these shares, and at the end it is possible to reconstruct the result from the different shares without revealing information about the initial secrets. This is done by using Lagrange interpolation.\\
There is a huge family of protocols build on this approach, and it is still evolving every day. This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similar to SPDZ.\\

@@

\subsection{Adversary Model}
In cryptography, the adversary which is considered is normally a third party that wants to get information for his own gain or want to sabotage something because of his own interest. In SMPC it is expected that the adversary is not only participating as one party, but could could control multiple parties. That means a potential adversary can control between one and all but one of the parties which are participating. The difference to the normal cryptographic adversary model is, that the adversary is controlling parties that are participating in the scheme and not a third party that tries to intercept communication.\\
In the area of SMPC there is normally a differentiation between two types of adversaries.

\subsubsection{Honest but curious}
The honest but curious adversary tries to gain as much information as possible while obeying to the rules of the protocol. Hence, it is impossible to know that this party is controlled by a third entity. This adversary tries to retrieve some of the secret information from the other parties, by using all the information he gets during the execution of the protocol.

\subsection{Malicious adversary}
While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol. This type of adversary can send different inputs to manipulate the results he receives. Furthermore, he wants to interrupt the whole protocol. That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications

@@

\subsubsection{Access structures}
There are a lot of different use cases for SMPC, so of course the adversary model changes depending on the system, the security needs, and the amount of parties involved. it is important to differentiate between these different use cases, because the different needs have big implications for security measurements that have to be taken and, thus, for runtime complexity. SMPC should run as fast as possible, while still providing all the necessary security guarantees.\\
To describe which party has which rights, and how many parties can be corrupted, so that the security can still be guaranteed, access structures were introduced.
An access structure describes how many parties are needed to form a so called qualified set. A qualified set is a group of parties which can pool their information to extract secret information. An access structure is called \(Q_l \)-access structure if at least \(l+1\) parties are needed to  form a qualified set. 
An \((n,t)\)-threshold scheme with \(t<\frac{n}{l}\) means that at least \(t+1\) parties are needed to form a qualified set. Here, \(n\) is the number of total parties, and \(t\) is number of possible adversaries while not revealing any secrets.\\
A full threshold scheme means that all parties are needed to form a qualified set. This implies, that, if all other parties except one are corrupted by an adversary, this one party can still rely on the fact that it secret remains secret. Of course this has a huge impact on overhead and performance, but in some cases it can provide needed security guarantees.
An access structure also does not have to be symmetric. In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is, for example, possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
Choosing the right access structure for the right use case is really important, because it has a realy big impact on performance and on the desired security level.

@@

\subsubsection{Overview}

@@

\subsubsection{MAMBA}
For this paper the more interesting part is the algorithmic part. But still, the importance of the compiler shouldn't be undervalued. Because while the algorthmic foundations might be more important from a theoretical point of view, the compiler, written in C++, enables the user to really use these technologies in a system where he can rely on the security gurantees given by the algorithms. The programs can be written for MAMBA in a syntax similiar to python's. In these programs, the user can specify which security settings he wishes to use. He can specify, how he wants his SMPC to exactly be executed, and which security parameters he wants to use. MAMBA is the compiler which then translates the instruction to bytecode and executes them. MAMBA is the part that gives the system the possibility to be properly used and tested. It makes it easy, to change the parameters, to be able to execute comparable tests, and to have a abstraction level for the user.\\

\subsubsection{Secret sharing}

@@

\subsubsection{Revealing secrets}

@@

\subsubsection{Offline phase}

There are two main things done in the offline phase. The first one is the generation of a global MAC key $\alpha$. This MAC key is later used, to authenticate the different shares at the end of the computation. This key is also distributively generated, i.e. only if all parties pool together their shares they can recover its value.\\

The second and and most important part in the offline phase, is the generation of the beaver triples. These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties. For each multiplication, one of the triples is needed.

@@

\subsubsection{Online Phase}

@@

\subsubsection{Upcoming changes to SCALE-MAMBA}
SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further. An important aspect is the usability in the practice. Because online and offline phase are integrated into each other, the complete runtime sometimes takes a lot of time. This is especially the case in full threshold access structures. Because of the integration, it is right now not possible to execute the offline phase earlier. One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem. The developer team tries to give MAMBA the possibility to compile code just in time as mentioned in [TODO: link to documentation]. That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and then in-time compiled and executed, without having to wait for the expensive offline calculations.\\

@@

\subsubsection{Outlook}
Their are a lot of future applications for Secure Multiparty Computation. With the threoretical research of the early years much more turning to praxis oriented research, todays SMPC protocols, and especially the successors of SPDZ, have a bright future. With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, the list will only grow more and more with the performance of the protocols and the same time the available computation power improving. \\
With SCALE-MAMBA being one of the first performant systems that can be really used in a real time scenario, it will probably not take a lot of more years until there will be proper commercial solutions. And if big companies start to adapt and use these theoretical foundations, then, in my opinion, SMPC will be a really important tool in computer science and especially in distrbuted environments. There are a lot more possible applications that could be coming in the next years, for example in the financial branche SMPC could play a huge role.\\
SMPC can be a technology that shapes the future, or it can be just a stepstone for a new technology to emerge. But definetly, privacy and security plays already a huge role in our society, and this will not diminish in any close future.

@@

\subsubsection{Conclusion}

>>>>

\section{Motivation and Introduction}\label{sec:intromoti}
In today's world, and even more in the future, there exist a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all of it. Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive or it has a high business value.\\
A basic example is an anonymous voting. In most elections there is a third party that has to aggregate and count the votes. For instance, in Germany's elections[TODO: LINK] this is done by specific people counting a small amount of votes and then forwarding the information to a central institution. The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time. So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of voters.\\
Another use case is the exchange of medical data. Medical data of patients is highly confidential and should never be published. But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it would be helpful to compare the symptoms to the diagnoses of other doctors.\\
Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
The basic problem is the evaluation of a common function, to which multiple parties have different inputs, without revealing these inputs. Although there are a lot of theoretical solutions for this problem, real-world implementations are still not performing on the desired level. So in the last years the research shifted more into the area of how to improve the performance of existing solutions while still guaranteeing a desired level of security.\\
This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven. This system is one of the forerunners in SMPC and combines a lot of research by different teams to provide a platform with highly performant SMPC. Nevertheless, this paper will at first give a short introduction into the history of SMPC and will try to explain its basic technologies to the reader. It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially taking a look on which different requirements are fulfilled by each. Ultimately, it will try to give an outlook what might be coming in the next years.\\

@@

\subsection{Beginnings}
As we saw in the introduction, there are a lot current and much more upcoming use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret. The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yao's millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their respective wealth). \\
These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
\subsection{Problem statement}
Formally, the SMPC problem can be stated as follows:
\begin{itemize}
\item Given a finite field \( \mathcal{F}_p=\{0,...,p-1\}\) with a prime p
\item Given \(n\) parties \(P_1,...,P_n\) participating in the computation
\item Given \(m\) secret inputs \(x_1,...,x_m\) with each of them hold by one of the parties
\item Given a function \(f(x_1,...,x_m)\)
\item Goal: Evaluation of the function \(f(x_1,...,x_m)\) while guaranteeing secrecy and correctness
\end{itemize}
\textbf{Secrecy:} If party \(P_i\) holds secret \(x_j\), then at the end of the evaluation no information regarding \(x_j\) is revealed to the other parties.\\
\textbf{Correctness:} At the end of the computation, the result is correct.\\\\
In a lot of cases each party has exactly one corresponding secret input, but this is not necessarily required. There can also be parties without input and parties with multiple inputs.\\
Now we will take a short look on two of the first protocols, which both are the basis for a whole family of SMPC schemes.
\subsection{Yao's Garbled Circuit}
There are typically two families, into which most of the SMPC protocols can be categorized. One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}. The basic idea is that, in a two party setting, one party, Alice, garbles the Boolean circuit that has to be evaluated and sends it to the second party, Bob. Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an encrypted way. At the end they exchange the results and can compute the result this way. 
\subsection{Shamirs Secret Sharing Scheme}
This paper will not further discuss the family of garbled circuit protocols, but rather the second family which are protocols based on so called Linear Secret Sharing. 
These type of protocols were introduced by Adi Shamir at a similar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
The basic idea of this approach is the the so called share-compute-reveal paradigm, which splits the evaluation of the function into three parts.\\
\textbf{Share:} First, the inputs secrets are split into multiple parts which then are shared shared with the other parties. Hence, every party holds a share of each input secret.\\
\textbf{Compute:} The parties evaluate the function with the secret input shares. The protocol has to be defined in a way, that the final result which each parties computes with the input shares is a respective share of the complete final resut.\\
\textbf{Reveal:} The parties agregate their shares of the final result and compute the final result together.\\
This paradigm already shows a lot of challenges arising for a concrete protocol, for instance:
\begin{itemize}
\item How can the secrets be shared in a homomorphic way, so that, after computing on shared input, the complete result can be recovered from the result shares?
\item While this can be done for addition in easy ways, how can this be done for multiplication as well?
\item How can the correctness of the computation be verified?
\item How can the overhead generated by a protocol can be kept in a reasonable size, so that a protocol is applicable to a real-world use case?
\end{itemize}
In Shamir's scheme the secret sharing is implemented by expressing a secret $s$ and its shares as a random polynomial $f(x)$, so that $f(0)=s$ and each share $s_i$ has the value $s_i=f(i)$. Computations are executed on these shares, and at the end it is possible to reconstruct the result from the different shares without revealing information about the initial secrets. This is done by using Lagrange interpolation. . The degree $p$ of the polynomial can be seen as a security parameter, because $p$ shares are needed to reconstruct the secret.\\
There is a huge family of protocols build on this approach, and it is still evolving every day. This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similar to SPDZ. \\

@@

\subsection{Adversary Model}
In cryptography, the adversary which is considered is normally a third party that wants to get information for his own gain or want to sabotage something because of his own interest. In SMPC it is expected that the adversary is not only participating as one party, but could could control multiple parties. That means a potential adversary can control between one and all but one of the parties which are participating. The difference to the normal cryptographic adversary model is, that the adversary is controlling parties that are participating in the scheme and not a third party that tries to intercept communication.\\
In the area of SMPC there is normally a differentiation between two types of adversaries.
\subsubsection{Honest but curious}
The honest but curious adversary tries to gain as much information as possible while obeying to the rules of the protocol. Hence, it is impossible to know that this party is controlled by a third entity. This adversary tries to retrieve some of the secret information from the other parties, by using all the information he gets during the execution of the protocol.
\subsubsectionx{Malicious adversary}
While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol. This type of adversary can send different inputs to manipulate the results he receives. Furthermore, he wants to interrupt the whole protocol. That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications

@@

\subsection{Access structures}
There are a lot of different use cases for SMPC, so of course the adversary model changes depending on the system, the security needs, and the amount of parties involved. it is important to differentiate between these different use cases, because the different needs have big implications for security measurements that have to be taken and, thus, for runtime complexity. SMPC should run as fast as possible, while still providing all the necessary security guarantees.\\
To describe which party has which rights, and how many parties can be corrupted, so that the security can still be guaranteed, access structures were introduced.
An access structure describes how many parties are needed to form a so called qualified set. A qualified set is a group of parties which can pool their information to extract secret information. An access structure is called \(Q_l \)-access structure if at least \(l+1\) parties are needed to  form a qualified set. 
An \((n,t)\)-threshold scheme with \(t<\frac{n}{l}\) means that at least \(t+1\) parties are needed to form a qualified set. Here, \(n\) is the number of total parties, and \(t\) is number of possible adversaries while not revealing any secrets.\\
A full threshold scheme means that all parties are needed to form a qualified set. This implies, that, if all other parties except one are corrupted by an adversary, this one party can still rely on the fact that it secret remains secret. Of course this has a huge impact on overhead and performance, but in some cases it can provide needed security guarantees.
An access structure also does not have to be symmetric. In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is, for example, possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
Choosing the right access structure for the right use case is really important, because it has a realy big impact on performance and on the desired security level.

@@

\subsection{Overview}

@@

\subsection{MAMBA}
For this paper the more interesting part is the algorithmic part. But still, the importance of the compiler shouldn't be undervalued. Because while the algorthmic foundations might be more important from a theoretical point of view, the compiler, written in C++, enables the user to really use these technologies in a system where he can rely on the security gurantees given by the algorithms. The programs can be written for MAMBA in a syntax similiar to python's. In these programs, the user can specify which security settings he wishes to use. He can specify, how he wants his SMPC to exactly be executed, and which security parameters he wants to use. MAMBA is the compiler which then translates the instruction to bytecode and executes them. MAMBA is the part that gives the system the possibility to be properly used and tested. It makes it easy, to change the parameters, to be able to execute comparable tests, and to have a abstraction level for the user.\\

\subsection{Secret sharing}

@@

\subsection{Revealing secrets}

@@

\subsection{Offline phase}

There are two main things done in the offline phase. The first one is the generation of a global Message Authentification Code(MAC) key $\alpha$. This MAC key is later used, to authenticate the different shares at the end of the computation.\\
The second and and most important part in the offline phase, is the generation of the beaver triples. These triples  are independent of the computation that is later performed in the online phase, but are used in the online phase to express multiplications as a combination of additions and multiplications with constants. For each multiplication, one of the triples is needed.

\subsubsection{Generation of the MAC key}
Message Authentification Codes (MACs) are used to guarantee the integrity and authenticity of a message. Basically, a MAC is a key that is used to generate a tag for a message. This tag is used to verify the identity of the sender and that the content of the message was not changed.\\
For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
\begin{itemize}
\item The MAC algorithm \(\text{MAC}(K,m)\rightarrow t\) takes a key \(K\) and a message \(m\) and outputs a tag \(t\)\\
\item The verficiation algorithm \(\text{VER}(K,m,t)\rightarrow 0/1\) takes a key \(K\), a message \(m\) and a tag \(t\) and outputs accept(\(1\)) if the tag is correct or reject(\(0\))\\ if it is not
\end{itemize}

The MAC key is distributively generated, i.e. only if all parties pool together their shares they can recover its value. 
[TODO Expand on MAC generation

@@

\subsection{Online Phase}

@@

\subsection{Upcoming changes to SCALE-MAMBA}
SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further. An important aspect is the usability in the practice. Because online and offline phase are integrated into each other, the complete runtime sometimes takes a lot of time. This is especially the case in full threshold access structures. Because of the integration, it is right now not possible to execute the offline phase earlier. One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem. The developer team tries to give MAMBA the possibility to compile code just in time as mentioned in [TODO: link to documentation]. That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and then in-time compiled and executed, without having to wait for the expensive offline calculations.\\

@@

\subsection{Outlook}
Their are a lot of future applications for Secure Multiparty Computation. With the threoretical research of the early years much more turning to praxis oriented research, todays SMPC protocols, and especially the successors of SPDZ, have a bright future. With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, the list will only grow more and more with the performance of the protocols and the same time the available computation power improving. \\
With SCALE-MAMBA being one of the first performant systems that can be really used in a real time scenario, it will probably not take a lot of more years until there will be proper commercial solutions. And if big companies start to adapt and use these theoretical foundations, then, in my opinion, SMPC will be a really important tool in computer science and especially in distrbuted environments. There are a lot more possible applications that could be coming in the next years, for example in the financial branche SMPC could play a huge role.\\
SMPC can be a technology that shapes the future, or it can be just a stepstone for a new technology to emerge. But definetly, privacy and security plays already a huge role in our society, and this will not diminish in any close future.

@@

\subsection{Conclusion}
